---
title: "RMark_Covariates"
output: html_document
date: "2024-09-27"
---

```{r}

library(dplyr)
library(RMark)


```

## Data Formatting

Resolve inconsistencies in frog_sex column. For frogs with multiple sexes listed over multiple captures, use the most recent observation. If the most recent observation is blank, use the most common sex identification.

```{r}

# Load the full capture data set
capture_data_full <- read.csv("Capture_Data.csv")

# Select relevant columns, remove duplicates, and arrange by PIT tag
capture_data_full <- capture_data_full %>% 
  select(Date, PIT_TagCode, frog_sex) %>% 
  unique() %>% 
  arrange(PIT_TagCode)

# Clean up the Date column (remove time and convert to Date format)
capture_data_full$Date <- gsub("00:00:00 EDT", "", capture_data_full$Date)
capture_data_full$Date <- as.Date(capture_data_full$Date, format = "%a %b %d %Y")

# Write function to resolve the frog)_sex, taking the most recent, or the most common non-empty sex identification if the most recent is empty
resolve_sex <- function(data) {
  # Check if there's any non-empty sex data
  non_empty_data <- data %>% filter(frog_sex != "")
  
  if (nrow(non_empty_data) == 0) {
    return(NA)  # Return NA if no valid sex is found
  }
  
  # Find the most recent sex
  most_recent <- data %>% filter(Date == max(Date)) %>% pull(frog_sex)
  
  if (most_recent != "") {
    return(most_recent)
  } else {
    # If most recent sex is empty, return the most common non-empty sex
    return(non_empty_data %>% 
             count(frog_sex) %>% 
             slice_max(n, n = 1) %>% 
             pull(frog_sex))
  }
}

# Apply the function to resolve the sex for each frog
resolved_sex_data <- capture_data_full %>%
  group_by(PIT_TagCode) %>%
  group_modify(~ mutate(.x, Sex_resolved = resolve_sex(.x))) %>%
  ungroup() %>% 
  select(PIT_TagCode, Sex_resolved) %>% 
  unique() %>% 
  rename(sex = Sex_resolved)%>%
  mutate(is_Female = ifelse(sex == "M", 0, ifelse(sex == "F", 1, NA))) %>% 
  filter(!is.na(sex) & sex != "U")

```

## Read in binary capture data formatted for RMark use and join it with sex data. 

```{r}

capture <- read.csv("Binary_Capture_History.csv", colClasses = c('numeric', 'character'))

capture <- right_join(capture, resolved_sex_data, 'PIT_TagCode')
capture$sex <- as.factor(capture$sex)

```

## Run model with sex covariate

Time Intervals

```{r}

# Intervals in days between primary periods
time.intervals.1 <- c(0,0,33,0,0,54,0,0,271,0,0,32,0,0,55,0,0,299,0,0,25,0,0,26,0,0,320,0,0,26,0,0,26,0,0,278,0,0,19,0,0,26,0,0,26,0)

# Months between primary periods
time.intervals.4 <- time.intervals.1/30

getwd()

```

Prep data and run model

```{r}

data_temp <- process.data(data = capture,
                  model = "Robust",
                  time.intervals = time.intervals.4,
                  groups = c("sex")
                 )

design_data_temp <- make.design.data(data_temp)


model_temp <- mark(data_temp, design_data_temp, model = "Robust",
              model.parameters = list(
                S = list(formula = ~time + sex),  
                p = list(formula = ~session),
                c = list(formula = ~session),
                GammaPrime = list(formula = ~time),
                GammaDoublePrime = list(formula = ~time)
              ))

N_temp <- model_temp$results$derived$`N Population Size`

survival_temp <- model_temp$results$real[1:30, ]

capture_estimates <- model_temp$results$real[grep("^p ", rownames(model_temp$results$real)), ]

recapture_estimates <- model_temp$results$real[grep("^c ", rownames(model_temp$results$real)), ]

?process.data()

```


```